---
title: "Coastal vulnerability"

output:
  html_document

knit: (function(input, encoding) {
    rmarkdown::render(
      input = input,
      encoding = encoding,
      envir = globalenv()
    )
  })
---

Identify coastal communities

```{r}
#rebuild and load the package
devtools::load_all() # restarts and loads

# List of packages for session
.packages = c("import")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
#lapply(.packages, require, character.only=TRUE)

#specifies what functions to import
# https://cran.r-project.org/web/packages/import/vignettes/import.html
import::from(dplyr, mutate, distinct, select, group_by, summarise, keep_when = filter)



```

```{r}
#set data dir
input.dir <- "~/coastalvulnerability-inputs"

```


```{r read indicators, eval=FALSE, include=FALSE}
sheetlink = "https://docs.google.com/spreadsheets/d/1j9GJKPjKgpf4w3bDeS2c8uamN8-L79I1BINjpQe0lCk/edit#gid=1817952998"
# 
tabla.susc <- read_googlesheet(sheetlink, sheetname = "relacion_datos_procesados_LOC", coltypes = "cccccccccccccccccccc")
tabla.susc <- read_googlesheet(sheetlink, sheetname = "resumen_indicadores_exposicion", coltypes = "cccccccccccccccccccc")


for(i in 1:nrow(tabla.susc)){
  
  this.susc <- tabla.susc[i,]
  
}

```


```{r extract communities}

#coordinates
crs.proj.utm <- "+proj=utm +zone=12 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +datum=NAD83"
crs.proj.wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

#Gulf of California and Yucatan Peninsula states
ents.selecc <- c("['Baja California Sur']", "['Baja California']", "['Sonora']", "['Sinaloa']", "['Nayarit']", "['Jalisco']", "['Yucatán']", "['Campeche']", "['Quintana Roo']")
ents.coast.selecc <- c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco", "Yucatán", "Campeche", "Quintana Roo")

#all coastal states
#ents.coast <- c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco", "Colima", "Michoacán de Ocampo", "Guerrero", "Oaxaca", "Chiapas", "Quintana Roo", "Yucatán", "Campeche", "Tabasco", "Veracruz de Ignacio de la Llave", "Tamaulipas")

#read county (municipio) shape and subset selected states
municipios.selecc <- sf::st_read(paste0(input.dir, "/shp_files/georef-mexico-municipality-millesime.shp")) %>% 
keep_when(sta_name %in% ents.selecc) 

#read atlantic ocean shp
atlantic.shp <- sf::st_read(paste0(input.dir,"/shp_files/atlantic_ocean.shp")) 
#read pacific ocean shp
pacific.shp <- sf::st_read(paste0(input.dir,"/shp_files/pacific_ocean.shp")) 

#read shp Mexico states
est.mex <- sf::st_read(paste0(input.dir,"/datos_INEGI_2020/dest2019gw.shp"))
estados.shp <- sf::st_read(paste0(input.dir,"/shp_files/ESTADOS.shp"))

#2020 Census data 
inegi.2020 <- readr::read_csv(paste0(input.dir,"/datos_INEGI_2020/conjunto_de_datos/conjunto_de_datos_iter_00_cpv2020.csv")) %>% 
  keep_when(LONGITUD!=0) %>% 
  keep_when(NOM_ENT %in% ents.coast.selecc)

#extract and save codes for states
ents.selecc.codes <- estados.shp %>% 
  sf::st_drop_geometry() %>% 
  keep_when(NOM_ENT %in% ents.coast.selecc)

ents.selecc.codes %>% 
readr::write_csv(paste0(input.dir,"/processed_data/edos_gcyuc.csv"))

#estados.mx <- cbind(est.2019, st_coordinates(st_centroid(est.2019)))

#select counties in selected states that are coastal
#pacific 
mun.cost.pac <- municipios.selecc %>% 
   mutate(pac_distance=sf::st_distance(., pacific.shp)) %>% 
   keep_when((as.numeric(pac_distance)) < 10000)
#atlantic
mun.cost.atl <- municipios.selecc %>% 
   mutate(atl_distance=sf::st_distance(., atlantic.shp)) %>% 
   keep_when((as.numeric(atl_distance)) < 10000)

#join counties from apcific and atlantic
mun.cost <- mun.cost.atl %>% 
  dplyr::bind_rows(mun.cost.pac)

#get codes for states and target municipalities
mun.ent.codes <- mun.cost %>% 
   sf::st_drop_geometry() %>% 
  mutate(mun_cve = gsub("\\[","",mun_code), 
         mun_cve = gsub("\\]","",mun_cve),
         mun_cve = gsub("\\'","",mun_cve)) %>% 
  mutate(mun_nam = gsub("\\[","",mun_name), 
         mun_nam = gsub("\\]","",mun_nam),
         mun_nam = gsub("\\'","",mun_nam)) %>% 
  mutate(ent_nam = gsub("\\[","",sta_name), 
         ent_nam = gsub("\\]","",ent_nam),
         ent_nam = gsub("\\'","",ent_nam)) %>% 
   mutate(ent_cve = gsub("\\[","",sta_code), 
         ent_cve = gsub("\\]","",ent_cve),
         ent_cve = gsub("\\'","",ent_cve)) %>% 
  select(mun_cve,mun_nam, ent_cve, ent_nam)

mun.ent.codes %>% readr::write_csv(paste0(input.dir,"/processed_data/","mun_ent_codes.csv"))


#get community data, output is spatial data layer
inegi.cost <- get_community_data(inegi.2020, crs.proj.utm, crs.proj.wgs, mun.cost, ents.coast.selecc)

inegi.cost.locs <- inegi.cost %>% 
  select(CVE_ENT:POBMAS)

sf::st_write(inegi.cost.locs, paste0(input.dir,"/processed_data/shp_files/","inegi_comm_cost.shp"), delete_layer = TRUE)


inegi.cost.buffer <-  inegi.cost %>% 
  select(id_row:POBTOT) %>% 
  sf::st_buffer(dist=10000)

sf::st_write(inegi.cost.buffer, paste0(input.dir,"/shp_files/","inegi_comm_buffer.shp"), delete_layer = TRUE)

inegi.cost.buffer.50 <-  inegi.cost %>% 
  select(id_row:POBTOT) %>% 
  sf::st_buffer(dist=50000)

sf::st_write(inegi.cost.buffer.50, paste0(input.dir,"/shp_files/","inegi_comm_buffer_50k.shp"), delete_layer = TRUE)

inegi.cost.comm <- inegi.cost %>% 
  sf::st_drop_geometry() %>% 
  select(CVE_ENT:NOM_LOC, POBTOT, deci_lon, deci_lat) %>% 
  distinct(CVE_ENT, CVE_LOC, CVE_MUN, NOM_ENT, NOM_MUN, NOM_LOC, POBTOT, deci_lon, deci_lat)

#get location codes for target locations

inegi.cost.comm %>% readr::write_csv(here::here("outputs","inegi_coast_comm.csv"))
inegi.cost.comm %>% readr::write_csv(paste0(input.dir,"/processed_data/","inegi_coast_comm.csv"))

#save full census data for selected communities

inegi.cost.pop <- inegi.cost %>% 
  sf::st_drop_geometry()

inegi.cost.pop %>% readr::write_csv(paste0(input.dir,"/processed_data/","inegi_coast_census.csv"))


sf::st_write(mun.cost, here::here("outputs","shp_files","mun_cost.shp"), delete_layer = TRUE)

make_map(mun.cost, inegi.cost, estados.shp, crs.proj.utm, thistitle="Total population",thisoutput=here::here("outputs","figures","map_locs_en.png"))

make_map(mun.cost, inegi.cost, estados.shp, crs.proj.utm, thistitle="Población total",thisoutput=here::here("outputs","figures","map_locs_sp.png"))


```

Download DENUE data economic units registered in Mexico by economic activity https://en.www.inegi.org.mx/app/mapa/denue/
Data updated November 2023
```{r get denue data, echo=FALSE}

inegi.comms  <-  readr::read_csv(here::here("outputs","inegi_coast_comm.csv")) 

inegi.comm.codes  <-  inegi.comms %>% 
  dplyr::pull(CVE_LOC)

denue.links= "links_DENUE_descarga_masiva_2023.csv"

unidades.denue <- readr::read_csv(paste0(input.dir,"/", denue.links))

#only run once to download, unzip files and aggregate data

#actividad.denue <- get_denue(unidades.denue, input.dir) 
#unzip_denue(input.dir)

#denue.inegi <- list.files(path=paste0(input.dir,"/denue/conjunto_de_datos"), pattern="denue_inegi_*.*")
#denue.inegi.full <- list.files(path=paste0(input.dir,"/denue/conjunto_de_datos"), pattern="denue_inegi_*.*", full.names = TRUE)

#list.denue <- 1:length(denue.inegi.full)

#lapply(list.denue, extract_denue, denue.inegi, denue.inegi.full, input.dir, inegi.comm.codes, inegi.comms)

denue.cost.comms <- list.files(path=paste0(input.dir,"/processed_data/denue"), pattern="denue_*.*", full.names = TRUE) 

denue.all <- lapply(denue.cost.comms, combine_denue)

denue.all.comm <- denue.all %>% 
  dplyr::bind_rows() %>% 
  dplyr::rename(CVE_ENT = ENTIDAD) %>% 
  make_code()

denue.all.comm %>% readr::write_csv(paste0(input.dir,"/processed_data/adaptive_capacity/","denue_all_ecactiv.csv"))


# remove fisheries
denue.other.act <- denue.cost.comms[!denue.cost.comms %in% c("/home/atlantis/coastalvulnerability-inputs/processed_data/denue/denue_fisheries_ecactiv.csv")]  

denue.other <- lapply(denue.other.act, combine_denue)
  
denue.other.comm <- denue.other %>% 
  dplyr::bind_rows() %>% 
  dplyr::rename(CVE_ENT = ENTIDAD) %>% 
  make_code()

denue.other.comm %>% readr::write_csv(paste0(input.dir,"/processed_data/adaptive_capacity/","denue_other_ecactiv.csv"))

denue.other.tot <- denue.other.comm %>% 
  group_by(NOM_ENT,NOM_MUN, CVE_MUN, NOM_LOC, CVE_LOC) %>% 
  summarise(tot_denue = sum(neg_tot))

readr::write_csv(denue.other.tot, paste0(input.dir,"/processed_data/adaptive_capacity/","denue_tot_ecactiv.csv"))

#schools and other educational resources  
denue.educa <- denue.other.comm %>% 
  keep_when(grepl("educación", nombre_act, ignore.case = TRUE) | grepl("educativos", nombre_act, ignore.case = TRUE) | grepl("Escuelas", nombre_act, ignore.case = TRUE) | grepl("Biblioteca", nombre_act, ignore.case = TRUE))
 
denue.educa.tot <- denue.educa %>% 
  mutate(activity = "education") %>% 
  group_by(CVE_MUN, CVE_LOC, activity) %>% 
  summarise(tot_eddenue = sum(neg_tot)) %>% 
  dplyr::left_join(inegi.comms, by =c("CVE_MUN", "CVE_LOC")) %>% 
  mutate(edu_pcapita = tot_eddenue/ POBTOT)

readr::write_csv(denue.educa.tot, paste0(input.dir,"/processed_data/adaptive_capacity/","denue_education_ecactiv.csv"))

#fishery and aquaculture businesses
denue.fish <- readr::read_csv("/home/atlantis/coastalvulnerability-inputs/processed_data/denue/denue_fisheries_ecactiv.csv")

denue.fish.tot <- denue.fish %>% 
  group_by(CVE_MUN, CVE_LOC) %>% 
  summarise(tot_denue = sum(neg_tot)) 

readr::write_csv(denue.fish.tot, paste0(input.dir,"/processed_data/adaptive_capacity/","denue_fishaqtot_ecactiv.csv"))

aqua.act <- c("Camaronicultura","Piscicultura y otra acuicultura, excepto camaronicultura")			

fishery.act <- c("Pesca de camarón","Pesca de sardina y anchoveta","Pesca de túnidos","Pesca y captura de peces, crustáceos, moluscos y otras especies")
				
denue.fish.num <- denue.fish %>% 
  keep_when(nombre_act %in% fishery.act) %>% 
  mutate(activity = "fisheries economic units") %>% 
  group_by(CVE_MUN, CVE_LOC, activity) %>% 
  summarise(tot_denue_fish = sum(neg_tot))  %>% 
  dplyr::left_join(inegi.comms, by =c("CVE_MUN", "CVE_LOC")) %>% 
  mutate(fish_pcapita = tot_denue_fish/ POBTOT) %>% 
  dplyr::left_join(denue.other.tot, by= c("NOM_ENT","NOM_MUN","CVE_MUN","NOM_LOC","CVE_LOC")) %>% 
  mutate(prop_uefish = tot_denue_fish / tot_denue) %>% 
  select(CVE_LOC, CVE_MUN, tot_denue_fish, fish_pcapita, prop_uefish) %>% 
  replace(is.na(.), 0) 

readr::write_csv(denue.fish.num, paste0(input.dir,"/processed_data/sensitivity/","denue_fisheries_ecactiv.csv"))

denue.aqua.num <- denue.fish %>% 
  keep_when(nombre_act %in% aqua.act) %>% 
  mutate(activity = "aquaculture economic units") %>% 
  group_by(CVE_MUN, CVE_LOC, COM_ID, activity) %>% 
  summarise(tot_denue_aqua = sum(neg_tot)) %>% 
  dplyr::left_join(inegi.comms, by =c("CVE_MUN", "CVE_LOC")) %>% 
  mutate(aqua_pcapita = tot_denue_aqua/ POBTOT) %>% 
  dplyr::left_join(denue.other.tot, by= c("NOM_ENT","NOM_MUN","CVE_MUN","NOM_LOC","CVE_LOC")) %>% 
  mutate(prop_ueaqua = tot_denue_aqua / tot_denue) %>% 
  select(CVE_LOC, CVE_MUN, tot_denue_aqua, aqua_pcapita, prop_ueaqua) %>% 
  replace(is.na(.), 0) 


readr::write_csv(denue.aqua.num, paste0(input.dir,"/processed_data/sensitivity/","denue_aqua_ecactiv.csv"))




```


```{r road access}

road.data <- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/road_access/","GACP_Grado de accesso por carretera_2018.xlsx"))

readr::write_csv(access.res, paste0(input.dir,"/processed_data/adaptive_capacity/road_access.csv"))



cat.tm <- road.data %>% 
  distinct(tm_centro)

cat.pav <- road.data %>% 
  distinct(dis_pav)

cat.gr <- road.data %>% 
  distinct(gr_access)

#not used data was from 1985
#note that OpenStreet Map data can be obtained from R package osmdata and could update this section
#https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html

# crs.proj.utm <- "+proj=utm +zone=12 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +datum=NAD83"
# crs.proj.wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
# 
# ext.carr.proj <- sf::st_read("~/coastalvulnerability-inputs/adaptive_capacity/road_access/carre1mgw.shp") %>% 
# sf::st_transform(., 6367)
# 
# inegi.cost.locs <- sf::st_read(paste0(input.dir,"/processed_data/shp_files/","inegi_comm_cost.shp"))
# 
# inegi.cost.locs.proj <- inegi.cost.locs %>% 
#   sf::st_transform(., 6367)
# 
# length.mt <- ext.carr.proj %>% 
#   sf::st_length() %>% 
#   sf::st_drop_geometry() %>% 
#   as.numeric() %>% 
#   tidyr::as_tibble() %>% 
#   dplyr::rename(length_m=value)
# 
# ext.carr.proj.dis <- ext.carr.proj %>% 
#   dplyr::bind_cols(length.mt)
# 
# #estimate distance to any road
# 
# inegi.carr.dist <- sf::st_distance(inegi.cost.locs.proj, ext.carr.proj.dis,  by_element = FALSE)
# 
# inegi.carr.dist.m <- inegi.carr.dist %>% 
#  apply(., 1, FUN = min)%>% 
#   tidyr::as_tibble() %>% 
#   mutate(dist_km = value/1000)
# 
# dist.table <- inegi.cost.locs %>% 
#   sf::st_drop_geometry() %>% 
#   dplyr::bind_cols(inegi.carr.dist.m) %>% 
#   select(CVE_LOC, CVE_MUN, dist_km)
#   
#   
# #estimate road density within community buffer
# 
# inegi.cost.buffer.proj <- sf::st_read(paste0(input.dir,"/shp_files/","inegi_comm_buffer.shp")) %>% 
# sf::st_transform(., 6367)
# 
# carretera.loc <- sf::st_intersection(ext.carr.proj.dis, inegi.cost.buffer.proj)
# 
# length.loc <- carretera.loc %>% 
#   sf::st_length() %>% 
#   sf::st_drop_geometry() %>% 
#   as.numeric() %>% 
#   tidyr::as_tibble() %>% 
#   dplyr::rename(length_loc=value)
# 
# ext.carr.proj.loc <- carretera.loc %>% 
#   dplyr::bind_cols(length.loc)
# 
# carretera.loc.sum <- tidyr::as_tibble(ext.carr.proj.loc) %>% 
#   sf::st_drop_geometry() %>% 
#   mutate(CVE_ENT = cve_ent, CVE_MUN = cve_mun, CVE_LOC = cve_loc) %>% 
#   make_code() %>% 
#   group_by(CVE_MUN, CVE_LOC) %>% 
#   summarise(den_loc_km = (sum(length_m)/1000)) %>% 
#   replace(is.na(.), 0) 
# 
# access.res <- dist.table %>% 
#   dplyr::left_join(carretera.loc.sum, by = c("CVELOC", "CVE_MUN"))



```

Employment centers
```{r}

emp.cent <- readxl::read_xlsx("~/coastalvulnerability-inputs/adaptive_capacity/employment_centers/cct_2021.xlsx") 

inegi.table <- "inegi_coast_comm.csv"
inegi.comms <- fix_inegi(inegi.table)

inegi.comm.codes  <-  inegi.comms %>% 
  dplyr::pull(CVE_LOC)

emp.cent.code <- emp.cent %>% 
  dplyr::rename(CVE_ENT= INMUEBLE_CV_ENT, NOM_ENT = INMUEBLE_C_NOM_ENT, CVE_MUN = INMUEBLE_CV_MUN, NOM_MUN= INMUEBLE_C_NOM_MUN, NOM_LOC=INMUEBLE_C_NOM_LOC) %>% 
  keep_when(NOM_ENT!="NA") %>% 
  keep_when(CVE_LOC %in% inegi.comm.codes) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII")) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII")) %>% 
  make_code()
  
emp.cent.code.sum <- emp.cent.code %>% 
  mutate(id_row = 1) %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(tot_emp = sum(id_row)) %>% 
  dplyr::left_join(inegi.comms, by= c("CVE_LOC", "CVE_MUN")) %>% 
  mutate(prop_emp = tot_emp/POBTOT) %>% 
  select(CVE_ENT, CVE_MUN, CVE_LOC, tot_emp, prop_emp) %>% 
  replace(is.na(.), 0) 

readr::write_csv(emp.cent.code.sum, paste0(input.dir,"/processed_data/adaptive_capacity/employment_centers.csv"))


```

Infrastructure indicators
```{r}

inf.ind <- readxl::read_xlsx("~/coastalvulnerability-inputs/adaptive_capacity/access_infrastructure/Base_indicadores_loc_PATP_abril_2023.xlsx") 

inegi.table <- "inegi_coast_comm.csv"
inegi.comms <- fix_inegi(inegi.table)

inegi.comm.codes  <-  inegi.comms %>% 
  dplyr::pull(CVE_LOC)

inf.ind.codes <- inf.ind %>% 
  mutate(CVE_MUN = `Clave municipio`, CVE_LOC = `Clave localidad`) %>% 
    select(CVE_MUN, CVE_LOC, r_1erN_auto, r_2doN_auto, r_3erN_auto, r_preescolar_auto, r_primaria_auto, r_secundaria_auto, r_CID) %>% 
  dplyr::rename(acc_1health=r_1erN_auto, acc_2health=r_2doN_auto, acc_3health=r_3erN_auto, acc_presc=r_preescolar_auto, acc_elem=r_primaria_auto, acc_sec = r_secundaria_auto, acc_dev=r_CID)

names.categ <- colnames(inf.ind.codes)
num.cols <- 3:length(names.categ)

new.tibble <- inf.ind.codes[,1:2]

new.inf.codes <- lapply(num.cols, add_varcategory, names.categ, inf.ind.codes)

#NAs correspond to not available values in the original table. These are mostly islands, where the corresponding centers are too far but distance was not measured, so we assigned the highest values

inf.table <- new.inf.codes %>% 
  dplyr::bind_cols(new.tibble, .) %>% 
  keep_when(CVE_LOC %in% inegi.comm.codes) %>% 
  replace(is.na(.), 0) 

inf.table.max <- inf.table %>%
  summarise(acc_1healthMax=max(acc_1health), acc_2healthMax=max(acc_2health), acc_3healthMax=max(acc_3health), acc_prescMax=max(acc_presc), acc_elemMax=max(acc_elem),
                   acc_secMax = max(acc_sec), acc_devMax=max(acc_dev))

inf.table.rev <- inf.table %>% 
  mutate(acc_1health = dplyr::if_else(acc_1health==0, inf.table.max$acc_1healthMax, acc_1health)) %>% 
  mutate(acc_2health = dplyr::if_else(acc_2health==0, inf.table.max$acc_2healthMax, acc_2health)) %>% 
  mutate(acc_3health = dplyr::if_else(acc_3health==0, inf.table.max$acc_3healthMax, acc_3health)) %>% 
  mutate(acc_presc = dplyr::if_else(acc_presc==0, inf.table.max$acc_prescMax, acc_presc)) %>% 
 mutate(acc_elem = dplyr::if_else(acc_elem==0, inf.table.max$acc_elemMax, acc_elem)) %>% 
 mutate(acc_sec = dplyr::if_else(acc_sec==0, inf.table.max$acc_secMax, acc_sec)) %>% 
 mutate(acc_dev = dplyr::if_else(acc_dev==0, inf.table.max$acc_devMax, acc_dev))


readr::write_csv(inf.table.rev, paste0(input.dir,"/processed_data/adaptive_capacity/access_infrastructure.csv"))

```

Poverty indicators and marginalization index
```{r}

inegi.table <- "inegi_coast_comm.csv"
inegi.comms <- fix_inegi(inegi.table)

inegi.comm.codes  <-  inegi.comms %>% 
  dplyr::pull(CVE_LOC)

pov.ind.1 <- readxl::read_xls("~/coastalvulnerability-inputs/adaptive_capacity/poverty_marginalization/IML_2020.xls", sheet = 2) 

pov.ind.comms.1 <- pov.ind.1 %>% 
  keep_when(CVE_LOC %in% inegi.comm.codes)

pov.ind.2 <- readxl::read_xls("~/coastalvulnerability-inputs/adaptive_capacity/poverty_marginalization/IML_2020.xls", sheet = 3) 

pov.ind.comms.2 <- pov.ind.2 %>% 
  keep_when(CVE_LOC %in% inegi.comm.codes)

pov.code <- pov.ind.comms.1 %>% 
  dplyr::bind_rows(pov.ind.comms.2) %>% 
  dplyr::rename(CVE_ENT= ENT, CVE_MUN = MUN) %>% 
  make_code()
  
pov.table <- pov.code %>% 
  select(CVE_LOC, CVE_MUN, ANALF:IM_2020)

#Social gap index

social.ind.1 <- readxl::read_xlsx("~/coastalvulnerability-inputs/adaptive_capacity/poverty_marginalization/IRS_localidades_2020.xlsx", sheet = 1) 
social.ind.2 <- readxl::read_xlsx("~/coastalvulnerability-inputs/adaptive_capacity/poverty_marginalization/IRS_localidades_2020.xlsx", sheet = 2) 

soc.ind.comms.1 <- social.ind.1 %>% 
  keep_when(cve_loc %in% inegi.comm.codes) %>% 
  dplyr::rename(CVE_LOC = cve_loc)

soc.ind.comms.2 <- social.ind.2 %>% 
  keep_when(cve_loc %in% inegi.comm.codes) %>% 
  dplyr::rename(CVE_LOC = cve_loc)

soc.code <- soc.ind.comms.1 %>% 
  dplyr::bind_rows(soc.ind.comms.2) %>% 
  make_code()

soc.table <- soc.code %>% 
  select(CVE_LOC, CVE_MUN, social_indx)

pov.res <- soc.table %>% 
  dplyr::left_join(pov.table, by = c("CVE_LOC", "CVE_MUN"))

readr::write_csv(pov.res, paste0(input.dir,"/processed_data/adaptive_capacity/poverty_indices.csv"))


```


Fisheries subsidies

```{r fisheries subsidies, echo=TRUE, message=TRUE, warning=FALSE}

inegi.table <- "inegi_coast_comm.csv"
inegi.cost.comm <- fix_inegi(inegi.table)

cost.codes.ent <- inegi.cost.comm %>%
  distinct(CVE_ENT, NOM_ENT) 

cost.codes.mun <- inegi.cost.comm %>%
  distinct(CVE_ENT, NOM_MUN, CVE_MUN) 

cost.codes.loc <- inegi.cost.comm %>%
   distinct(NOM_ENT, CVE_ENT, NOM_MUN, CVE_MUN, NOM_LOC, CVE_LOC) 
 

#googlesheet to keep track of names that need correction
googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1eJXjwlHqEzSAZaTmE_XyN93fu6EI0I8igRH4AdvXOl8")

correct.ent <- locs.pesca %>% 
  select(-NOM_LOC_REV)

correct.loc <- locs.pesca %>% 
  select(-NOM_ENT_REV) %>% 
  keep_when(!is.na(NOM_LOC_REV))

#gas subsidies
gas.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/padron_beneficiarios_combustible_2011-2019.xlsx")) %>% 
   keep_when(año>=2014) %>% 
  mutate(localidad=gsub("-"," ",localidad)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = municipio, id = "Latin-ASCII")) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = municipio, id = "Latin-ASCII")) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = localidad, id = "Latin-ASCII")) %>% 
  keep_when(tipo_cobertura=="LOCALIDAD") %>% 
  extract_comms(inegi.cost.comm) %>% 
  mutate(NOM_LOC = dplyr::if_else(grepl("BAHIA TORTUGAS", unidadeconmica), "BAHIA TORTUGAS", 
                                         dplyr::if_else(grepl("SCPP EMANCIPACION SC DE RL", unidadeconmica), "BAHIA TORTUGAS", NOM_LOC)))

gas.subsidy.comms <-  gas.subsidy %>% 
  dplyr::left_join(correct.ent, by =c("NOM_LOC","NOM_ENT")) %>% 
  mutate(NOM_ENT = dplyr::if_else(!is.na(NOM_ENT_REV), NOM_ENT_REV, NOM_ENT)) %>%
  dplyr::rename(ent_old = entidad) %>% 
  dplyr::left_join(correct.loc, by =c("NOM_LOC","NOM_ENT")) %>% 
  mutate(NOM_LOC = dplyr::if_else(!is.na(NOM_LOC_REV), NOM_LOC_REV, NOM_LOC)) %>%
  dplyr::rename(mun_old = municipio, loc_old = localidad) %>% 
  select(NOM_LOC, NOM_ENT, beneficiario, monto_conapesca) %>% 
  keep_when(monto_conapesca>0)

#some states, municipalites and locations are wrong so starting from ENT
gas.subsidy.ent <- gas.subsidy.comms %>% 
  dplyr::left_join(cost.codes.ent, by=c("NOM_ENT")) %>% 
  select(CVE_ENT, dplyr::everything())

gas.subsidy.loc <- gas.subsidy.ent %>% 
   dplyr::left_join(inegi.cost.comm, by=c("CVE_ENT","NOM_LOC","NOM_ENT")) %>% 
  select(CVE_LOC, CVE_MUN, CVE_ENT, dplyr::everything())

gas.subsidy.codes <- gas.subsidy.loc %>% 
  keep_when(!is.na(CVE_LOC)) %>% 
  mutate(id_row = 1) %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_bengas=sum(id_row), tot_bengas=sum(monto_conapesca)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(bengas_perc = (tot_bengas / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_bengas, bengas_perc)


#BIENPESCA

bienpesca.files <- c(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/bien_pescadores_y_acuicultores_2020.xlsx"),
                     paste0(input.dir,"/adaptive_capacity/fishery_subsidies/bienpesca_2021.xlsx"),
                     paste0(input.dir,"/adaptive_capacity/fishery_subsidies/bienpesca_2022.xlsx"),
                     paste0(input.dir,"/adaptive_capacity/fishery_subsidies/bienpesca_2023.xlsx"))

bpesca.subsidy.comms <- lapply(bienpesca.files, clean_bienpesca, inegi.cost.comm) %>% 
  dplyr::bind_rows()

bpesca.subsidy.codes <- bpesca.subsidy.comms %>% 
  extract_comms(inegi.cost.comm) %>% 
  select(-NOM_MUN, -CVE_MUN, -cve_mun) %>% 
  dplyr::left_join(inegi.cost.comm, by = c("NOM_ENT","CVE_ENT","NOM_LOC","CVE_LOC","POBTOT")) %>%  
  make_code()
 
bpesca.subsidy.res <- bpesca.subsidy.codes %>% 
  mutate(id_row = 1) %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_bpesca=sum(id_row), tot_bpesca=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(bpesca_perc = (tot_bpesca / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_bpesca, bpesca_perc)

#subsidies to aquaculture to purchase seed
seed.files <- c(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/semilla_acuicola_2021.xlsx"),
                     paste0(input.dir,"/adaptive_capacity/fishery_subsidies/semilla_acuicola_2022.xlsx"),
                     paste0(input.dir,"/adaptive_capacity/fishery_subsidies/semilla_acuicola_2023.xlsx"))

seed.subsidy.comms <- lapply(seed.files, clean_seedfiles) %>% 
  dplyr::bind_rows()

seed.subsidy.codes <- seed.subsidy.comms %>% 
   extract_comms(inegi.cost.comm) %>% 
   make_code()

seed.subsidy.res <- seed.subsidy.codes %>% 
  mutate(id_row = 1) %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_seed=sum(id_row), tot_seed=sum(`MONTO FEDERAL`)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(seed_perc = (tot_seed / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_seed, seed_perc)

#subsidies to fund genetic line improvements
gen.files <- c(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/recursos_geneticos_acuicolas_2020.xlsx"),
               paste0(input.dir,"/adaptive_capacity/fishery_subsidies/lineas_geneticas_mejoradas_2021.xlsx"),
               paste0(input.dir,"/adaptive_capacity/fishery_subsidies/lineas_geneticas_mejoradas_2022.xlsx"),
               paste0(input.dir,"/adaptive_capacity/fishery_subsidies/lineas_geneticas_mejoradas_2023.xlsx"))

gen.subsidy.comms <- lapply(gen.files, clean_seedfiles) %>% 
  dplyr::bind_rows()

gen.subsidy.codes <- gen.subsidy.comms %>% 
   extract_comms(inegi.cost.comm) %>% 
   make_code()

gen.subsidy.res <- gen.subsidy.codes %>% 
  mutate(id_row = 1) %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_gen=sum(id_row), tot_gen=sum(`MONTO FEDERAL`)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(gen_perc = (tot_gen / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_gen, gen_perc)


#vessel modernization

boat.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_modernizacion_barcos_2011-2019.xlsx"))

boat.subsidy.comms <- extract_comms(boat.subsidy, inegi.cost.comm)

boat.subsidy.table <- boat.subsidy.comms %>% 
  keep_when(año>=2014) %>% 
  keep_when(tipo_cobertura=="LOCAL") %>% 
  select(-cve_loc) %>% 
  mutate(id_row = 1, monto_gob_edo = as.numeric(monto_gob_edo), monto_productor = as.numeric(monto_productor), mont_tot = (monto_productor + monto_gob_edo + monto_conapesca) )%>% 
  keep_when(mont_tot!=0) %>% 
  select(-NOM_MUN) %>% 
  dplyr::left_join(inegi.cost.comm, by = c("NOM_ENT","NOM_LOC")) %>%  
  make_code() 

boat.subsidy.codes <- boat.subsidy.table %>%
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_boat=sum(id_row), tot_boat=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(boat_perc = (tot_boat / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_boat, boat_perc)

#subsidies for product commercialization
commerce.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_comercializacion_2014-2019.xlsx"))

commerce.subsidy.comms <- extract_comms(commerce.subsidy, inegi.cost.comm)

commerce.subsidy.table <- commerce.subsidy.comms %>% 
  keep_when(tipo_cobertura=="LOCAL") %>% 
  select(-cve_loc) %>% 
  mutate(mont_tot = (monto_gob_edo + monto_conapesca))%>% 
  keep_when(mont_tot!=0) %>% 
   make_code() 
  
commerce.subsidy.codes <- commerce.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_comm=sum(no_beneficiarios), tot_comm=sum(mont_tot)) %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(comm_perc = (tot_comm / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_comm, comm_perc)


#subsidies for aquaculture
aqua.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_acuacultura_2014-2019.xlsx"))
#this file does not have project characteristics for 2019 data, that is in aquaculture_incentives_2019.xlsx

aqua.subsidy.comms <- extract_comms(aqua.subsidy, inegi.cost.comm)

aqua.subsidy.table <- aqua.subsidy.comms %>% 
  keep_when(tipo_cobertura=="LOCAL") %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  make_code()

aqua.subsidy.codes <- aqua.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_aqua=sum(id_row), tot_aqua=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(aqua_perc = (tot_aqua / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_aqua, aqua_perc)


#subsidies to support seafood consumption
consump.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_consumo_2014-2019.xlsx"))

consump.subsidy.comms <- extract_comms(consump.subsidy, inegi.cost.comm)

consump.subsidy.table <- consump.subsidy.comms %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  make_code()

consump.subsidy.codes <- consump.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_cons=sum(id_row), tot_cons=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(cons_perc = (tot_cons / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_cons, cons_perc)

#subsidies to support fishery value chains
value.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_cadenas_productivas_2011-2019.xlsx")) %>% 
  keep_when(año>=2014)

value.subsidy.comms <- extract_comms(value.subsidy, inegi.cost.comm)

value.subsidy.table <- value.subsidy.comms %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  make_code() 

value.subsidy.codes <- value.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_value=sum(id_row), tot_value=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(value_perc = (tot_value / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_value, value_perc)


#subsidies to strengthen fisheries
str.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_fortalecimiento_2011-2019.xlsx")) %>% 
  keep_when(año>=2014)

str.subsidy.comms <- extract_comms(str.subsidy, inegi.cost.comm)

str.subsidy.table <- str.subsidy.comms %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  make_code() 

str.subsidy.codes <- str.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_str=sum(id_row), tot_str=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(str_perc = (tot_str / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_str, str_perc)

#subsidies to strengthen inspection and surveillance
ins.subsidy <- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_vigilancia_2011-2019.xlsx")) %>% 
  keep_when(año>=2014) %>% 
  dplyr::rename(localidad = localidad_ue, municipio = municipio_ue)

ins.subsidy.comms <- extract_comms(ins.subsidy, inegi.cost.comm)

ins.subsidy.table <- ins.subsidy.comms %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  make_code() 

ins.subsidy.codes <- ins.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_ins=sum(id_row), tot_ins=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(ins_perc = (tot_ins / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_ins, ins_perc)


#subsidies to strengthen the legal framework
leg.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_ordenamiento_2011-2019.xlsx")) %>% 
  keep_when(año>=2014) 

leg.subsidy.comms <- extract_comms(leg.subsidy, inegi.cost.comm)

leg.subsidy.table <- leg.subsidy.comms %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  dplyr::left_join(correct.loc, by =c("NOM_LOC","NOM_ENT")) %>% 
  mutate(NOM_LOC = dplyr::if_else(!is.na(NOM_LOC_REV), NOM_LOC_REV, NOM_LOC)) %>%
  make_code()

leg.subsidy.codes <- leg.subsidy.table %>% 
  keep_when(!is.na(CVE_LOC)) %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_leg=sum(id_row), tot_leg=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(leg_perc = (tot_leg / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_leg, leg_perc)


#subsidies to reduce effort in industrial fishing
eff.subsidy<- readxl::read_xlsx(paste0(input.dir,"/adaptive_capacity/fishery_subsidies/beneficiarios_reduccion_esfuerzo_2011-2018.xlsx")) %>% 
  keep_when(año>=2014)

eff.subsidy.comms <- extract_comms(eff.subsidy, inegi.cost.comm)

eff.subsidy.table <- eff.subsidy.comms %>% 
  mutate(id_row = 1, mont_tot = monto_conapesca)%>% 
  keep_when(mont_tot!=0) %>% 
  make_code()

eff.subsidy.codes <- eff.subsidy.table %>% 
  group_by(CVE_LOC, CVE_MUN) %>% 
  summarise(no_eff=sum(id_row), tot_eff=sum(mont_tot)) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(inegi.cost.comm, by = c("CVE_LOC","CVE_MUN")) %>% 
  mutate(eff_perc = (tot_eff / POBTOT)) %>% 
  select(CVE_LOC, CVE_MUN, no_eff, eff_perc)


all.fishery.subsidies <- inegi.cost.comm %>% 
  dplyr::left_join(bpesca.subsidy.res, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(seed.subsidy.res, by = c("CVE_LOC", "CVE_MUN")) %>%
  dplyr::left_join(gas.subsidy.codes,by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(boat.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(commerce.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(aqua.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(consump.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(value.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(str.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(ins.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(leg.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(eff.subsidy.codes, by = c("CVE_LOC", "CVE_MUN")) %>% 
  dplyr::left_join(gen.subsidy.res, by = c("CVE_LOC", "CVE_MUN"))

fishery.tot.subs <- all.fishery.subsidies %>% 
  replace(is.na(.), 0) 

fishery.comm.subs <- fishery.tot.subs %>%
  mutate(sum = rowSums(dplyr::across(no_bpesca:gen_perc), na.rm=TRUE)) %>% 
  keep_when(sum>0)

#fishery.tot.subs %>% summarise(across(no_bpesca:gen_perc, sum)) #use to test that all columns have value

fishery.tot.subs %>% distinct(CVE_LOC)

readr::write_csv(fishery.tot.subs, paste0(input.dir,"/processed_data/adaptive_capacity","fisheries_subsidies.csv"))

```

Extract permit information from the Registro Nacional de Pesca y Acuacultura RNPA 
https://www.gob.mx/conapesca/documentos/registro-nacional-de-pesca-y-acuacultura-rnpa

```{r get catch, echo=TRUE, message=TRUE, warning=TRUE}

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco", "Yucatán", "Campeche", "Quintana Roo"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#only communities considered in the analysis
inegi.cost.comm  <- readr::read_csv(paste0(input.dir,"/processed_data/","inegi_coast_comm.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC))

inegi.cost.mun <- inegi.cost.comm %>% 
  distinct(NOM_MUN, CVE_MUN)

permit.files <- list.files(path = paste0(input.dir,"/sensitivity/fisheries_data"), full.names = TRUE)
no.permitfiles <- 1:length(permit.files)

permit.data.inegi <- lapply(no.permitfiles, extract_permits, permit.files, inegi.cost.comm, coast.selecc)

permit.table <- permit.data.inegi %>% 
  dplyr::bind_rows() %>% 
  keep_when(!is.na(ENTIDAD)) %>% 
  distinct(NOM_ENT, rnp_code, NOM_MUN, NOM_LOC, ENTIDAD, MUN, LOC, CVE_MUN, CVE_LOC, COM_ID, deci_lat, deci_lon, POBTOT) %>% 
  select(-NOM_MUN) %>% 
  dplyr::left_join(inegi.cost.mun, by = "CVE_MUN")

permit.per.loc <- permit.table %>% 
  mutate(index_no = 1) %>% 
  group_by(NOM_ENT, NOM_MUN, NOM_LOC, ENTIDAD, MUN, LOC, CVE_MUN, CVE_LOC, COM_ID, deci_lat, deci_lon, POBTOT) %>% 
  summarise(tot_perm=sum(index_no)) %>% 
  mutate(perm_pcapita = tot_perm/POBTOT) %>% 
  select(CVE_MUN, CVE_LOC, tot_perm, perm_pcapita)

readr::write_csv(permit.per.loc, paste0(input.dir,"/processed_data/sensitivity/fishery_permits.csv"))

catch.files <- list.files(path = paste0(input.dir,"/catch"), full.names = TRUE)
no.catchfiles <- 1:length(catch.files)

#match to rnpa number

catch.rnpa <- lapply(no.catchfiles,get_rnpa_catch, catch.files, coast.selecc) 

rnpa.catch.table <- catch.rnpa %>%
  dplyr::bind_rows() %>% 
  mutate(rnp_code = as.character(rnp_code))

catch.rnpa.codes <- rnpa.catch.table %>%
  distinct(rnp_code) 

catch.permit.table <- catch.rnpa.codes %>% 
  dplyr::left_join(permit.table, by = c("rnp_code"))

catch.loc <- catch.permit.table %>% 
  dplyr::left_join(rnpa.catch.table, by = c("rnp_code")) %>% 
  keep_when(!is.na(ENTIDAD)) %>% 
  select(-NOM_MUN) %>% 
  dplyr::left_join(inegi.cost.mun, by = "CVE_MUN") %>% 
  group_by(NOM_ENT,NOM_MUN,CVE_MUN, NOM_LOC, CVE_LOC, COM_ID, deci_lat, deci_lon,POBTOT) %>% 
  summarise(tot_catch_mt = sum(tot_weight_kg)/ 1000) %>% 
  mutate(catch_percapita = tot_catch_mt/ POBTOT) %>% 
  select(CVE_MUN, CVE_LOC, tot_catch_mt, catch_percapita)

readr::write_csv(catch.loc, paste0(input.dir,"/processed_data/sensitivity/fishery_catch.csv"))


```



Assign catch to locations

```{r}
ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco", "Yucatán", "Campeche", "Quintana Roo"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")
coast.selecc.pac <- c("BAJA CALIFORNIA SUR","BAJA CALIFORNIA","SONORA","SINALOA","NAYARIT","JALISCO")

loc.pacifico <- readxl::read_xlsx(paste0(input.dir,"/Localidades_pesqueras_Pacifico.xlsx")) %>% 
  dplyr::rename(NOM_ENT= ESTADO, NOM_LOC = LOCALIDAD_REVISADA) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  select(-LOCALIDAD_ORIGINAL, -ID) %>% 
  keep_when(NOM_ENT %in% coast.selecc.pac)

#all coastal communities, including those with no census data
inegi.cost.comm  <- readr::read_csv(paste0(input.dir,"/processed_data/","all_coastal_communities_INEGI.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC))

catch.files <- list.files(path = paste0(input.dir,"/catch"), full.names = TRUE)
no.catchfiles <- 1:length(catch.files)

catch.locations <- lapply(no.catchfiles, extract_locations, catch.files, coast.selecc)
  
catch.locs.all <- catch.locations %>% 
  dplyr::bind_rows() %>% 
  distinct(NOM_ENT, NOM_LOC)

 catch.locs.inegi <-  catch.locs.all %>% 
    mutate(NOM_LOC = gsub(" \\(B\\.C\\.S\\.\\)","", NOM_LOC)) %>%
    mutate(NOM_LOC=dplyr::if_else(NOM_LOC=="BAHIA KINO","BAHIA DE KINO",
                                         dplyr::if_else(NOM_LOC=="PTO ADOLFO LOPEZ MATEOS","PUERTO ADOLFO LOPEZ MATEOS",
                                                 dplyr::if_else(NOM_LOC=="PUERTO VIEJO (EN SAN CARLOS)","PUERTO SAN CARLOS",
                                                                dplyr::if_else(NOM_LOC=="PUNTA FINAL (B. DE LOS ANGELES)","BAHIA DE LOS ANGELES",
                                                                               dplyr::if_else(NOM_LOC=="AGUAS LITORALES (BAHIA TORTUGAS)","BAHIA TORTUGAS", 
                                                                                              dplyr::if_else(NOM_LOC=="YAVAROS", "YAVAROS (ISLA LAS VIEJAS)", NOM_LOC))))))) %>% 
    mutate(NOM_ENT=dplyr::if_else(NOM_LOC=="TOPOLOBAMPO","SINALOA", NOM_ENT)) %>% 
    distinct(NOM_ENT, NOM_LOC) 
 
 catch.locs.inegi.all <- catch.locs.inegi %>% 
 dplyr::left_join(inegi.cost.comm, by = c("NOM_ENT","NOM_LOC")) %>% 
    group_by(NOM_ENT, NOM_LOC) %>%
    mutate(dupe = dplyr::n()>1)
  
 catch.locs.inegi.all %>% 
     keep_when(is.na(ENTIDAD)) %>% 
    write_csv(paste0(input.dir,"/processed_data/localidades_captura_inegi.csv"))
 
  #subset locations in the pacific and match with landing sites
  catch.locs.pac <- catch.locs.inegi %>% 
    keep_when(is.na(ENTIDAD)) %>% 
    keep_when(NOM_ENT %in% coast.selecc.pac) %>% 
    dplyr::left_join(loc.pacifico, by = c("NOM_ENT","NOM_LOC"))
     
  #used to find duplicates https://stackoverflow.com/questions/6986657/find-duplicated-rows-based-on-2-columns-in-data-frame-in-r
  #there are multiple landing codes by location, so not using
  # catch.file.sites.table %>% 
  #   group_by(NOM_ENT, NOM_LOC) %>%
  #   mutate(dupe = dplyr::n()>1)
  # 
  
  
   
  
```


Natural protected areas

```{r}
crs.proj.utm <- "+proj=utm +zone=12 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +datum=NAD83"
crs.proj.wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

prot.area <- sf::st_read("~/coastalvulnerability-inputs/adaptive_capacity/natural_protected_areas/226_ANP_ITRF08_27022024.shp") %>% 
sf::st_transform(., 6367)

#estimate road access within community buffer

inegi.cost.buffer.proj <- sf::st_read(paste0(input.dir,"/shp_files/","inegi_comm_buffer_50k.shp")) %>% 
sf::st_transform(., 6367) %>% 
  select(id_row:deci_ln, geometry)

prot.area.buff <- sf::st_intersection(prot.area, inegi.cost.buffer.proj)

area.int <- prot.area.buff %>% 
  sf::st_area() %>% 
  sf::st_drop_geometry() %>% 
  as.numeric() %>% 
  tidyr::as_tibble() %>% 
  dplyr::rename(area_loc_m2=value)

ext.area.proj.loc <- prot.area.buff %>% 
  dplyr::bind_cols(area.int)

area.loc.sum <- tidyr::as_tibble(ext.area.proj.loc) %>% 
  sf::st_drop_geometry() %>% 
  mutate(CVE_ENT = cve_ent, CVE_MUN= cve_mun, CVE_LOC = cve_loc) %>% 
  make_code() %>% 
  group_by(CVE_ENT, CVE_MUN, CVE_LOC) %>% 
  summarise(anp_area_loc_kmsq = (sum(area_loc_m2)/1e+6)) %>% 
  replace(is.na(.), 0) 


readr::write_csv(area.loc.sum, paste0(input.dir,"/processed_data/adaptive_capacity/natural_protected_area_buffer.csv"))

```

Numero de unidades acuicolas
Hectáreas de cultivo

```{r}

inegi.cost.buffer.proj <- sf::st_read(paste0(input.dir,"/shp_files/","inegi_comm_buffer.shp")) %>% 
 sf::st_transform(., 6367) %>% 
  select(id_row:deci_ln, geometry)

aquaculture.units <- sf::st_read(paste0(input.dir,"/sensitivity/aquaculture_polygons/","aquaculture_polygons_2023.shp")) 

ua.aqua.proj <- aquaculture.units %>% 
  sf::st_transform(., 6367)

#fixed geometries 
#https://github.com/r-spatial/sf/issues/870
ua.area.buff <-sf::st_intersection(sf::st_make_valid(ua.aqua.proj), inegi.cost.buffer.proj)

area.int <- ua.area.buff %>% 
  sf::st_area() %>% 
  sf::st_drop_geometry() %>% 
  as.numeric() %>% 
  tidyr::as_tibble() %>% 
  dplyr::rename(area_loc_m2=value)

ext.area.proj.loc <- ua.area.buff %>% 
  dplyr::bind_cols(area.int)

uarea.loc.sum <- tidyr::as_tibble(ext.area.proj.loc) %>% 
  sf::st_drop_geometry() %>% 
  mutate(CVE_ENT = cve_ent, CVE_MUN= cve_mun, CVE_LOC = cve_loc) %>% 
  make_code() %>% 
  group_by(CVE_ENT, CVE_MUN, CVE_LOC) %>% 
  summarise(ua_area_loc_kmsq = (sum(area_loc_m2)/1e+6)) %>% 
  replace(is.na(.), 0) 


readr::write_csv(uarea.loc.sum, paste0(input.dir,"/processed_data/sensitivity/aquaculture_units_buffer.csv"))


```


Adaptive capacity from 2020 census data

```{r}

inegi.table <- "inegi_coast_comm.csv"
inegi.comms <- fix_inegi(inegi.table)

inegi.comm.codes  <-  inegi.comms %>% 
  dplyr::pull(CVE_LOC)


inegi.data  <- readr::read_csv(paste0(input.dir,"/processed_data/","inegi_coast_census.csv"))
  
cost.data.pob <- inegi.data %>% 
  select(-POBTOT) %>% 
  make_code()


#Infrastructure

# VPH_INTER Private dwelling with Internet
# VPH_CEL Private dwelling with cell phone
# VPH_TIC Private dwellings with information infrastructure: TOTHOG - VPH_SINTIC Private dwellings without information infrastructure
# VPH_C_SERV Private dwellings with electricity, water tubing inside or outside the home, but within the property, as well as sewer
# VPH_AEASP Private dwellings with water tubing and public water service

infra.com.in <- cost.data.pob %>% 
  dplyr::select(CVE_MUN, CVE_LOC, VPH_INTER, VPH_CEL, VPH_SINTIC, VPH_C_SERV, VPH_AEASP, TOTHOG) %>% 
  mutate(across(VPH_INTER:TOTHOG, as.numeric)) %>% 
  mutate(PVPH_INTER = VPH_INTER/ TOTHOG) %>% 
  mutate(PVPH_CEL = VPH_CEL/ TOTHOG) %>%
  mutate(PVPH_TIC = (TOTHOG - VPH_SINTIC)/TOTHOG) %>%
  mutate(PVPH_C_SERV = VPH_C_SERV/ TOTHOG) %>%
  mutate(PVPH_AEASP = VPH_AEASP/ TOTHOG) %>% 
  keep_when(!is.na(VPH_INTER)) %>% 
  select(CVE_MUN, CVE_LOC, PVPH_INTER:PVPH_AEASP)


# Distancia a la vía más cercana
# Distancia a la carretera mas cercana

dist.via <- read_xlsx(here("inputs","adaptive_capacity","tabla_comunidades_carreteras.xlsx")) %>% 
  dplyr::select(COM_ID,NEAR_DIST_metros) %>% 
  dplyr::rename(dist_via = NEAR_DIST_metros) 

long.caminos <- read_xlsx(here("inputs","adaptive_capacity","carret_mpios_edos2.xlsx")) %>% 
  mutate(COM_MUN = paste(CVE_EDO,CVE_MUN, sep="_")) %>%
  group_by(COM_MUN) %>% 
  summarise(long_cam = sum(Length_km)) 

acceso.pav <- read_csv(here("inputs","adaptive_capacity","Grado de accesibilidad a carretera pavimentada a nivel localidad, 2010.csv")) %>%
  dplyr::select(COM_ID, distancia)

disp.agua <- st_read(here("inputs","adaptive_capacity","dispaguagw.shp")) 

disp.min <- calc_buffer(thispolygon=disp.agua,thisfield="Disp_HM3",inegi.cost.utm, thisfunction='min')

disp.loc <- inegi.cost.utm %>% 
  bind_cols(disp.min) %>% 
  as_tibble %>% 
  dplyr::select(COM_ID,layer) %>% 
  dplyr::rename(disp_agua = layer)


infra.com.in <- cost.data.pob %>% 
  dplyr::select(COM_ID, COM_MUN, NOM_LOC,TVIVPARHAB, VPH_AGUADV) %>% 
  mutate(across(TVIVPARHAB:VPH_AGUADV, as.numeric)) %>% 
  mutate(hogares_aguaent = VPH_AGUADV/TVIVPARHAB) %>% 
  left_join(long.caminos, by ="COM_MUN") %>% 
  left_join(dist.via, by ="COM_ID") %>% 
  left_join(acceso.pav, by = "COM_ID") %>% 
  left_join(disp.loc, by = "COM_ID") %>% 
  dplyr::select(COM_ID, NOM_LOC, hogares_aguaent, long_cam ,dist_via, distancia ,disp_agua) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(dist_via = -1 * dist_via) %>% # menor es positivo
  mutate(distancia = -1 * distancia) # menor es positivo


#caracteristicas de los hogares
carac.hogares.in <- cost.data.pob %>% 
  dplyr::select(CVE_LOC, CVE_MUN, POBTOT, VPH_INTER, 
                VPH_2YMASD, VPH_SINTIC, VPH_PISOTI, VPH_DRENAJ, TVIVPARHAB,VPH_SNBIEN,VPH_C_ELEC) %>% 
  mutate(across(VPH_2YMASD:VPH_C_ELEC, as.numeric)) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(hogares_recamaras = VPH_2YMASD/TVIVPARHAB) %>% #+ positivo
  mutate(hogares_tecnologia = ((TVIVPARHAB - VPH_SINTIC)/TVIVPARHAB)) %>% #+ positivo 
  mutate(hogares_ptierra = 1-(VPH_PISOTI/TVIVPARHAB)) %>% #- positivo
  mutate(hogares_servicios = ((TVIVPARHAB - VPH_SNBIEN)/TVIVPARHAB)) %>% #+ positivo 
  mutate(hogares_eelectrica = (VPH_C_ELEC/TVIVPARHAB)) %>% #+ positivo 
  mutate(hogares_drenaje = (VPH_DRENAJ/TVIVPARHAB)) %>% #+ positivo 
  dplyr::select(COM_ID, NOM_LOC, hogares_recamaras, hogares_tecnologia, hogares_ptierra,hogares_servicios,hogares_eelectrica, hogares_drenaje) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))


# Estructura del empleo
# Población económicamente activa	Población de 12 años y más económicamente activa		PEA
# Población empleada	Población de 12 años y más ocupada		POCUPADA
# Grado promedio de escolaridad	-		GRAPROES
# Población con servicios de salud	Población afiliada a servicios de salud		PDER_SS
# Población de 12 años y más	Personas de 12 a 130 años de edad.	P_12YMAS

comp.eco <- read_csv(here("inputs","adaptive_capacity","ECI.csv")) %>% 
  dplyr::select(COM_MUN,`Number of Employees Midpoint ECI`) %>% 
  dplyr::rename(in_compeco="Number of Employees Midpoint ECI")

ind.mar.in <- read_xls(here("inputs","adaptive_capacity","IML_2020.xls"), sheet =2) %>% 
  mutate(COM_ID = paste(as.numeric(ENT),as.numeric(MUN),as.numeric(LOC),sep="_")) %>% 
  dplyr::select(COM_ID,IM_2020) %>% 
  dplyr::rename(in_marg=IM_2020)

ind.mar <- read_xls(here("inputs","adaptive_capacity","IML_2020.xls"), sheet =3) %>% 
  mutate(COM_ID = paste(as.numeric(ENT),as.numeric(MUN),as.numeric(LOC),sep="_")) %>% 
  dplyr::select(COM_ID,IM_2020) %>% 
  dplyr::rename(in_marg=IM_2020) %>% 
  bind_rows(ind.mar.in)


carac.empleo.in <- cost.data.pob %>% 
  dplyr::select(COM_ID, COM_MUN, NOM_LOC, POBTOT, PEA, POCUPADA, GRAPROES, PDER_SS, P_12YMAS) %>% 
  mutate(across(PEA:P_12YMAS, as.numeric)) %>% 
  mutate(pob_ecactiva = PEA/P_12YMAS) %>% #+ positivo
  mutate(pob_empleada = POCUPADA/P_12YMAS) %>% #+ positivo
  mutate(grad_esc = GRAPROES) %>% #+ positivo
  mutate(pob_salud = PDER_SS/POBTOT) %>% #+ positivo 
  left_join(comp.eco, by="COM_MUN") %>% #+ positivo 
  left_join(ind.mar, by="COM_ID") %>% #+ positivo 
  dplyr::select(COM_ID, NOM_LOC, pob_ecactiva,pob_empleada,grad_esc,pob_salud, in_compeco, in_marg) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))
  
 

#poblacion indigena
#Población en hogares censales indígenas		PHOG_IND

#alfabetismo indigena por municipio
#Poblacion indígena alfabeta y analfabeta en hombres y mujeres de 15 y más. 
ana.ind <- st_read(here("inputs","adaptive_capacity","alfaindigw.shp")) %>% 
  as_tibble %>% 
  mutate(prop_inana =  pi15analfa/piobindi) %>% 
  na.omit %>% 
  mutate(COM_MUN = paste(as.numeric(CVE_ENT),as.numeric(CVE_MUN),sep="_")) %>% 
  dplyr::select(COM_MUN,prop_inana) %>% 
  distinct(COM_MUN,prop_inana)

#economia indigena

#Población activa, Población desocupada, Población inactiva, Población masculina, Población ocupada, 12 años
eco.ind <- st_read(here("inputs","adaptive_capacity","ecoindigw.shp")) %>% 
  as_tibble %>% 
  mutate(prop_indes =  pidesocup/pobindi) %>% 
  na.omit %>% 
  mutate(COM_MUN = paste(as.numeric(CVE_ENT),as.numeric(CVE_MUN),sep="_")) %>% 
  dplyr::select(COM_MUN,prop_indes) %>% 
  distinct(COM_MUN,prop_indes)

#migracion indigena

mig.ind <- st_read(here("inputs","adaptive_capacity","migindigw.shp")) %>% 
  as_tibble %>% 
  na.omit %>% 
  mutate(prop_inmig =  pilnacoten/pobindi) %>% 
  mutate(prop_inmig = if_else(is.nan(prop_inmig),0,prop_inmig)) %>% 
  mutate(COM_MUN = paste(as.numeric(CVE_ENT),as.numeric(CVE_MUN),sep="_")) %>% 
  dplyr::distinct(COM_MUN,prop_inmig) 

#Población total nacidos en la entidad, Población total nacidos en otra entidad, Población total nacidos no especificada, Población total que reside en la entidad y Población total que residen en otra entidad.

pob.indig.in <- cost.data.pob %>% 
  dplyr::select(COM_ID, COM_MUN, NOM_LOC, POBTOT, PHOG_IND) %>% 
  mutate(prop_pind = 1-(as.numeric(PHOG_IND)/POBTOT)) %>% #+ negativo
  left_join(mig.ind, by="COM_MUN") %>% #+ negativo
  left_join(ana.ind,by="COM_MUN") %>% #+ negativo
  left_join(eco.ind,by="COM_MUN") %>% #+ negativo
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  dplyr::select(COM_ID, NOM_LOC, prop_pind, prop_inmig, prop_inana, prop_indes) 


# Cambio en la población económicamente activa (2010-2020)	Población de 12 años y más económicamente activa		PEA
# Población que reside en otro estado	Población de 5 años y más residente en otra entidad en marzo de 2015		PRESOE15
# Cambio en la población soltera			P12YM_SOLT
# Cambio en los hogares femeninos	Hogares en viviendas particulares habitadas donde la persona de referencia es mujer.	HOGJEF_F
#Población sin religión o sin adscripción religiosa	Personas que declararon no tener religión o no estar adscritas en alguna.	PSIN_RELIG

censo.2010 <- fread(here("inputs","datos_INEGI_2010","iter_00_cpv2010","conjunto_de_datos","iter_00_cpv2010.csv")) %>% 
  as_tibble %>% 
  filter(!is.na(longitud)) %>% 
  mutate(COM_ID=paste(as.numeric(entidad),as.numeric(mun),as.numeric(loc),sep="_")) %>% 
   dplyr::select(COM_ID,pobtot, pea, p_12ymas, p12ym_solt,hogjef_f,psin_relig) %>% 
  mutate(across(pea:psin_relig, as.numeric)) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))

dis.personal.in <- cost.data.pob %>% 
  dplyr::select(COM_ID, NOM_LOC, POBTOT, PEA, P_5YMAS, PRESOE15, P12YM_SOLT, P_12YMAS, HOGJEF_F,PSIN_RELIG) %>% 
  mutate(across(PEA:PSIN_RELIG, as.numeric)) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  left_join(censo.2010, by="COM_ID") %>% 
  mutate(cam_pe = (pea/p_12ymas)-(PEA/P_12YMAS)) %>% #+ + positivo
  mutate(cam_mig = 1-(PRESOE15/P_5YMAS)) %>% #+ negativo
  mutate(cam_pobsol = (p12ym_solt/p_12ymas)-(P12YM_SOLT/P_12YMAS)) %>% #+ positivo
  mutate(cam_hogfem = (hogjef_f/pobtot)-(HOGJEF_F/POBTOT)) %>% #+ positivo
  mutate(cam_religion = (psin_relig)-(PSIN_RELIG/POBTOT)) %>% #+ positivo
  dplyr::select(COM_ID, NOM_LOC, cam_pe, cam_mig, cam_pobsol, cam_hogfem) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))
  

datasetname = "comp_pob"
comp.pob.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 

datasetname = "pobreza"
pobr.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 

datasetname = "carac_hogares"
carac.hogares.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 

datasetname = "infraestructura"
infra.com.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 
 
datasetname = "carac_empleo"
carac.empleo.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 


datasetname = "pob_indigena"
pob.indig.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 

 datasetname = "disrupcion_pers"
dis.personal.in %>%
     mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  pca_factor(., datasetname) 


adap.datos.com <- cost.data.pob %>% 
  dplyr::select(COM_ID,ENTIDAD,NOM_ENT,MUN,NOM_MUN,LOC,NOM_LOC,POBTOT) %>% 
  left_join(comp.pob.in, by=c("COM_ID", "NOM_LOC")) %>% 
  left_join(pobr.in, by=c("COM_ID", "NOM_LOC")) %>% 
  left_join(carac.hogares.in, by=c("COM_ID", "NOM_LOC")) %>% 
  left_join(carac.empleo.in, by=c("COM_ID", "NOM_LOC")) %>% 
  left_join(infra.com.in, by=c("COM_ID", "NOM_LOC")) %>% 
  left_join(pob.indig.in, by=c("COM_ID", "NOM_LOC")) %>% 
 left_join(dis.personal.in, by=c("COM_ID", "NOM_LOC")) %>% 
  mutate(NOM_LOC=tolower(stri_trans_general(str =NOM_LOC, id = "Latin-ASCII"))) %>% 
  mutate(NOM_MUN=tolower(stri_trans_general(str =NOM_MUN, id = "Latin-ASCII"))) %>% 
  mutate(NOM_ENT=tolower(stri_trans_general(str =NOM_ENT, id = "Latin-ASCII")))


adap.datos.com %>% write_csv(here("inputs","adaptive_capacity","data_summary","datos_capacidadadaptacion.csv"))

adap.datos.norm <- adap.datos.com %>%   
 dplyr::select(-ENTIDAD, -NOM_ENT, -MUN, -NOM_MUN, -LOC, -POBTOT) %>% 
    mutate_all(~replace(., is.na(.), 0)) 


```

